---
title: "2. Classical vs Bayesian Regression"
author: "Harm Schuett"
date: "12/10/2020"
output: html_document
---

```{r}
library(tidyverse)
library(broom)
library(patchwork)
library(ragg)
library(rstan)

theme_set(theme_bw() +
  theme(
    text = element_text(size = 9),
    panel.grid = element_blank()
  ))
```

## Creating simulated data and OLS regressions

```{r}
set.seed(888)
n_samples <- 50
n_obs <- 50
x <- rnorm(n = n_obs, 0, 1)
gen_data <- function(n) tibble(u = rnorm(n, 0, 20), x = x, y = 1 + 2 * x + u)

# gen samples
samples <- tibble(id = 1:n_samples)
samples$data <- map(rep.int(n_obs, n_samples), gen_data)
```




## Fig. 1: Classical Hypothesis tests

### OLS fits for 50 samples

```{r}
# gen ols estimates
samples$ols_fit <- map(samples$data, function(dat) tidy(lm(y ~ x, data = dat)))
ols_estimates <- samples %>%
  unnest(ols_fit) %>%
  mutate(param = if_else(term == "x", "a1", "a0")) %>%
  select(id, param, estimate) %>%
  spread(key = param, value = estimate) %>%
  mutate(
    plabs = paste0("(", round(a0, 1), ",", round(a1, 1), ")"),
    first_sample = as.factor(c(1, rep.int(0, times = n_samples - 1)))
  )
```

### OlS estimates for the first sample

```{r}
samples$ols_fit[[1]]
```

```{r}
range_x <- range(ols_estimates$a0)
range_y <- range(ols_estimates$a1)
nudge_x <- 0.1
sample_a0 <- ols_estimates$a0[1]
sample_a1 <- ols_estimates$a1[1]
r <- sd(ols_estimates$a1)
```


### Panel A code

```{r fig.height=3.5, fig.width=3.25, dpi=300}
panA.1 <-
  ols_estimates %>%
  ggplot(aes(x = a0, y = a1, shape = first_sample, color = first_sample)) +
  geom_point(size = 2) +
  xlim(range_x) +
  ylim(range_y) +
  scale_color_manual(values = c("gray80", "black")) +
  labs(
    x = expression(" Intercept estimate " ~ hat(a)[0]),
    y = expression(" Slope estimate " ~ hat(a)[1]),
    subtitle = "Sampling variation"
  ) +
  geom_segment(aes(color = first_sample), xend = 1, yend = 2, linetype = 2, alpha = 0.5) +
  annotate("point", x = 1, y = 2, color = "black", size = 2, shape = 15) +
  annotate("text", x = 1.4, y = 1.9, label = "Truth (1, 2)", hjust = 0, color = "black", size = 3) +
  annotate("text",
    x = sample_a0 + 0.4,
    y = sample_a1 + -0.1,
    label = paste0("Sample (", round(sample_a0, 1), ", ", round(sample_a1, 1), ")"),
    hjust = 0, color = "black", size = 3
  ) +
  annotate("point", x = 0, y = 0, color = "black", size = 2, shape = 15) +
  annotate("text", x = 0.4, y = -0.1, label = "H0 (0, 0)", hjust = 0, color = "black", size = 3) +
  theme(legend.position = "none")
panA.1
```

### Panel B code

```{r fig.height=3.5, fig.width=3.25, dpi=300}

panB.1 <- ols_estimates %>%
  ggplot(aes(x = a0, y = a1, shape = first_sample, color = first_sample)) +
  geom_point(size = 2) +
  xlim(range_x) +
  ylim(range_y) +
  scale_color_manual(values = c("gray80", "black")) +
  labs(
    x = expression(" Intercept estimate " ~ hat(a)[0]),
    y = expression(" Slope estimate " ~ hat(a)[1]),
    subtitle = "Distance from hypothesis"
  ) +
  theme(legend.position = "none") +
  geom_segment(aes(color = first_sample), xend = 0, yend = 0, linetype = 2, alpha = 0.5) +
  annotate("point", x = 0, y = 0, color = "black", size = 2, shape = 15) +
  annotate("text", x = 0.4, y = -0.1, label = "H0 (0, 0)", hjust = 0, color = "black", size = 3) +
  annotate("text",
    x = sample_a0 + 0.4,
    y = sample_a1 + -0.1,
    label = paste0("Sample (", round(sample_a0, 1), ", ", round(sample_a1, 1), ")"),
    hjust = 0, color = "black", size = 3
  ) +
  geom_hline(yintercept = c(r, -r), color = "black") +
  geom_segment(
    x = -4, xend = -4, y = -r, yend = r, color = "black",
    arrow = arrow(length = unit(0.07, "inches"), ends = "both", type = "closed")
  ) +
  annotate("text",
    x = -3.4, y = -2.2,
    size = 3, hjust = 0, vjust = 0, color = "black",
    label = "One standard deviation"
  )

panB.1
```

### Saving figure

```{r fig.height=3.5, fig.width=6.5, dpi=600}
fig1 <- panA.1 + panB.1 + plot_annotation(tag_levels = "A")
agg_png("../out/figs/fig1.png", width = 6.5, height = 3.5, units = "in", res = 600)
fig1
# ggsave(, fig1, device = agg_png, width = 6, height = 4, dpi = 600)
invisible(dev.off())
fig1
```


## Fitting Bayesian models

### Scaled data

```{r}
sample1 <- samples$data[[1]]

input_data <- list(
  y = scale(sample1$y)[, 1],
  x = scale(sample1$x)[, 1],
  N = nrow(sample1)
)

u_y <- mean(sample1$y)
u_x <- mean(sample1$x)
s_y <- sd(sample1$y)
s_x <- sd(sample1$x)
```

### Very weak prior version

```{r}
very_weak_regularization <- "
data{
  int<lower=1> N;
  real y[N];
  real x[N];
}
parameters{
  real a0;
  real a1;
  real<lower=0> sigma;
}
model{
  vector[N] mu;
  sigma ~ exponential( 1 );
  a0 ~ normal( 0 , 1 );
  a1 ~ normal( 0 , 1 );
  for ( i in 1:N ) {
    mu[i] = a0 + a1 * x[i];
  }
  y ~ normal( mu , sigma );
}
"
fit_1 <- stan(
  model_code = very_weak_regularization,
  data = input_data,
  iter = 5000,
  warmup = 2000,
  chains = 2,
  seed = 1234
)

posterior_draws_1 <- as.data.frame(as.matrix(fit_1))
print(fit_1, pars = c("sigma", "a0", "a1"))
```

#### Posterior descriptives for rescaled parameters

```{r}
rescaled_posterior_1 <- data.frame(
  a0 = posterior_draws_1$a0 * s_y + u_y - (posterior_draws_1$a1 * (s_y / s_x) * u_x),
  a1 = posterior_draws_1$a1 * (s_y / s_x)
)

summary(rescaled_posterior_1)
```

```{r}
quantile(rescaled_posterior_1$a1, probs = c(0.05, 0.95))
```

```{r}
emp_a1 <- ecdf(rescaled_posterior_1$a1)
emp_a1(0)
```


### Weakly informative version

```{r}
weakly_informative <- "
data{
  int<lower=1> N;
  real y[N];
  real x[N];
}
parameters{
  real a0;
  real a1;
  real<lower=0> sigma;
}
model{
  vector[N] mu;
  sigma ~ exponential( 1 );
  a0 ~ normal( 0 , 1 );
  a1 ~ normal( 0 , 0.2 );
  for ( i in 1:N ) {
    mu[i] = a0 + a1 * x[i];
  }
  y ~ normal( mu , sigma );
}
// generated quantities {
//   vector[N] log_lik;
//   for (n in 1:N) log_lik[n] = normal_lpdf(y[n] | a0 + a1 * x[n], sigma);
// }
"
fit_2 <- stan(
  model_code = weakly_informative,
  data = input_data,
  iter = 5000,
  warmup = 2000,
  chains = 2,
  seed = 1234
)

posterior_draws_2 <- as.data.frame(as.matrix(fit_2))
print(fit_2, pars = c("sigma", "a0", "a1"))
```

#### Posterior descriptives for rescaled parameters

```{r}
rescaled_posterior_2 <- data.frame(
  a0 = posterior_draws_2$a0 * s_y + u_y - (posterior_draws_2$a1 * (s_y / s_x) * u_x),
  a1 = posterior_draws_2$a1 * (s_y / s_x)
)

summary(rescaled_posterior_2)
```

```{r}
quantile(rescaled_posterior_2$a1, probs = c(0.05, 0.95))
```

```{r}
emp2_a1 <- ecdf(rescaled_posterior_2$a1)
emp2_a1(0)
```





## Fig. 2: Bayesian approach with very weak priors

### Data for a1 posterior panels

```{r}
# helper function for panel D
get_dens <- function(ppd, probs = c(0.05, 0.95)) {
  dens <- density(ppd)
  mode3 <- dens$x[dens$y == max(dens$y)]
  mode3.den <- dens$y[dens$x == mode3]
  dens.df <- data.frame(x = dens$x, y = dens$y)
  quantiles <- quantile(ppd, prob = probs)
  dens.df$quant <- factor(findInterval(dens.df$x, quantiles))
  dens.df$Q1 <- quantiles[1]
  dens.df$Q2 <- quantiles[2]
  return(dens.df)
}

post_dens <- get_dens(rescaled_posterior_1$a1)
```

### Data for prior panels

```{r}
x_length <- seq(-10, 10, .1)
x_length2 <- seq(0, 20, .1)
priors <- data.frame(
  a_den = dnorm(x = x_length, mean = 0, sd = 1),
  sigma_den = dexp(x = x_length2, rate = (1)),
  x1 = x_length,
  x2 = x_length2
)
```

### Panel A - D code

```{r}
range_a0 <- c(
  min(min(rescaled_posterior_1$a0), min(rescaled_posterior_2$a0)), 
  max(max(rescaled_posterior_1$a0), max(rescaled_posterior_2$a0))
)
range_a1 <- c(
  min(min(rescaled_posterior_1$a1), min(rescaled_posterior_2$a1)), 
  max(max(rescaled_posterior_1$a1), max(rescaled_posterior_2$a1))
)
```


```{r fig.height=3.5, fig.width=6.5, dpi=300}
panC.2 <-
  ggplot(data = rescaled_posterior_1, aes(x = a0, y = a1)) +
  stat_density_2d(aes(fill = ..level..), geom = "polygon", colour = "white") +
  annotate("point", x = 1, y = 2, color = "black", size = 2, shape = 15) +
  annotate("text", x = 1.4, y = 1.9, label = "Truth (1, 2)", hjust = 0, color = "black", size = 3) +
  annotate("point",
    x = mean(rescaled_posterior_1$a0), y = mean(rescaled_posterior_1$a1),
    color = "black", size = 2, shape = 17
  ) +
  annotate("text",
    x = -5, y = 10,
    label = paste0(
      "Posterior Mean\n(",
      round(mean(rescaled_posterior_1$a0), 1),
      ", ",
      round(mean(rescaled_posterior_1$a1), 1),
      ")"
    ),
    hjust = 0.5, color = "black", size = 3
  ) +
  annotate("segment",
    x = mean(rescaled_posterior_1$a0), xend = -5,
    y = mean(rescaled_posterior_1$a1), yend = 8.5
  ) +
  annotate("point", x = 0, y = 0, color = "black", size = 2, shape = 15) +
  annotate("text", x = 0.4, y = -0.1, label = "H0 (0, 0)", hjust = 0, color = "black", size = 3) +
  scale_x_continuous(expand = c(0, 0), limits = range_a0) +
  scale_y_continuous(expand = c(0, 0), limits = range_a1) +
  scale_fill_continuous(low = "grey95", high = "grey30") +
  theme(legend.position = "none") +
  labs(
    x = expression(a[0]),
    y = expression(a[1]),
    subtitle = quote("Posterior: Pr(" * a[0] * "," * a[1] * " | y, x)")
  )
panA.2 <-
  ggplot(data = priors, aes(x = x1, y = a_den)) +
  geom_area(color = "black", fill = "grey30", alpha = 0.5) +
  scale_y_continuous(expand = expansion(mult = c(0.001, 0.05))) +
  annotate("text", x = 9, y = 0.38, label = "N(0, 1)", hjust = 1, color = "black", size = 3) +
  scale_x_continuous(expand = c(0, 0)) +
  labs(
    y = "Density",
    x = expression(a[k]),
    subtitle = quote("Prior: Pr(" * a[k] * ")")
  )
panB.2 <-
  ggplot(data = priors, aes(x = x2, y = sigma_den)) +
  geom_area(color = "black", fill = "grey30", alpha = 0.5) +
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.05))) +
  scale_x_continuous(expand = c(0, 0)) +
  annotate("text", x = 19, y = 0.95, label = "Exp(1)", hjust = 1, color = "black", size = 3) +
  labs(
    y = "Density",
    x = expression(sigma),
    subtitle = quote("Prior: Pr(" * sigma * ")")
  )

panD.2 <-
  post_dens %>%
  ggplot(aes(x = x)) +
  geom_ribbon(aes(ymin = 0, ymax = y, fill = quant), alpha = 0.7) +
  geom_line(aes(y = y), color = "black") +
  labs(
    x = expression(a[1]),
    y = expression("Density"),
    subtitle = quote("Posterior: Pr(" * a[1] * " | y, x)")
  ) +
  scale_x_continuous(expand = c(0, 0), limits = range_a1) +
  scale_y_continuous(expand = expansion(mult = c(0.0, 0.05))) +
  scale_fill_manual(values = c("grey50", "grey70", "grey50")) +
  theme(legend.position = "None") +
  annotate("segment", x = 7, xend = 12, y = 0.02, yend = 0.05) +
  annotate("text", x = 11.5, y = 0.052, label = "90% Credible\ninterval", color = "black", size = 3, hjust = 0) +
  coord_flip()

layout <- "
ACCD
BCCD
"

fig2 <- panA.2 + panB.2 + panC.2 + panD.2 +
  plot_layout(design = layout) +
  plot_annotation(tag_levels = "A")
fig2
```

### Saving figure

```{r fig.height=3.5, fig.width=6.5, dpi=600}
agg_png("../out/figs/fig2.png", width = 6.5, height = 3.5, units = "in", res = 600)
fig2
invisible(dev.off())
```




## Fig. 3: Power of Priors



### Data for a1 posterior panels

```{r}
post_dens2 <- get_dens(rescaled_posterior_2$a1)
```

### Data for prior panels

```{r}
x_length <- seq(-10, 10, .1)
x_length2 <- seq(0, 20, .1)
priors2 <- data.frame(
  a_den = dnorm(x = x_length, mean = 0, sd = 0.5),
  sigma_den = dexp(x = x_length2, rate = (1)),
  x1 = x_length,
  x2 = x_length2
)
```

### Panel A - D code

```{r fig.height=3.5, fig.width=6.5, dpi=300}
range_bx <- range(rescaled_posterior_2$a0) #+ c(-0.1, +0.1)
range_by <- range(rescaled_posterior_2$a1) #+ c(-0.1, +0.1)

panC.3 <-
  ggplot(data = rescaled_posterior_2, aes(x = a0, y = a1)) +
  stat_density_2d(aes(fill = ..level..), geom = "polygon", colour = "white") +
  annotate("point", x = 1, y = 2, color = "black", size = 2, shape = 15) +
  annotate("text", x = 1.4, y = 1.9, label = "Truth (1, 2)", hjust = 0, color = "black", size = 3) +
  annotate("point",
    x = mean(rescaled_posterior_2$a0), y = mean(rescaled_posterior_2$a1),
    color = "black", size = 2, shape = 17
  ) +
  annotate("text",
    x = -5, y = 10,
    label = paste0(
      "Posterior Mean\n(",
      round(mean(rescaled_posterior_2$a0), 1),
      ", ",
      round(mean(rescaled_posterior_2$a1), 1),
      ")"
    ),
    hjust = 0.5, color = "black", size = 3
  ) +
  annotate("segment",
    x = mean(rescaled_posterior_2$a0), xend = -5,
    y = mean(rescaled_posterior_2$a1), yend = 8.5
  ) +
  annotate("point", x = 0, y = 0, color = "black", size = 2, shape = 15) +
  annotate("text", x = 0.4, y = -0.1, label = "H0 (0, 0)", hjust = 0, color = "black", size = 3) +
  scale_x_continuous(expand = c(0, 0), limits = range_a0) +
  scale_y_continuous(expand = c(0, 0), limits = range_a1) +
  scale_fill_continuous(low = "grey95", high = "grey30") +
  theme(legend.position = "none") +
  labs(
    x = expression(a[0]),
    y = expression(a[1]),
    subtitle = quote("Posterior: Pr(" * a[0] * "," * a[1] * " | y, x)")
  )
panA.3 <-
  ggplot(data = priors2, aes(x = x1, y = a_den)) +
  geom_area(color = "black", fill = "grey30", alpha = 0.5) +
  scale_y_continuous(expand = expansion(mult = c(0.001, 0.05))) +
  annotate("text", x = 9, y = 0.76, label = "N(0, 0.2)", hjust = 1, color = "black", size = 3) +
  scale_x_continuous(expand = c(0, 0)) +
  labs(
    y = "Density",
    x = expression(a[k]),
    subtitle = quote("Prior: Pr(" * a[1] * ")")
  )
panB.3 <-
  ggplot(data = priors2, aes(x = x2, y = sigma_den)) +
  geom_area(color = "black", fill = "grey30", alpha = 0.5) +
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.05))) +
  scale_x_continuous(expand = c(0, 0)) +
  annotate("text", x = 19, y = 0.95, label = "Exp(1)", hjust = 1, color = "black", size = 3) +
  labs(
    y = "Density",
    x = expression(sigma),
    subtitle = quote("Prior: Pr(" * sigma * ")")
  )

panD.3 <-
  post_dens2 %>%
  ggplot(aes(x = x)) +
  geom_ribbon(aes(ymin = 0, ymax = y, fill = quant), alpha = 0.7) +
  geom_line(aes(y = y), color = "black") +
  labs(
    x = expression(a[1]),
    y = expression("Density"),
    subtitle = quote("Posterior: Pr(" * a[1] * " | y, x)")
  ) +
  scale_x_continuous(expand = c(0, 0), limits = range_a1) +
  scale_y_continuous(expand = expansion(mult = c(0.0, 0.05))) +
  scale_fill_manual(values = c("grey50", "grey70", "grey50")) +
  theme(legend.position = "None") +
  annotate("segment", x = 7, xend = 12, y = 0.02, yend = 0.05) +
  annotate("text", x = 11.5, y = 0.052, label = "90% Credible\ninterval", color = "black", size = 3, hjust = 0) +
  coord_flip()

layout <- "
ACCD
BCCD
"

fig3 <- panA.3 + panB.3 + panC.3 + panD.3 +
  plot_layout(design = layout) +
  plot_annotation(tag_levels = "A")
fig3
```

### Saving figure

```{r fig.height=3.5, fig.width=6.5, dpi=600}
agg_png("../out/figs/fig3.png", width = 6.5, height = 3.5, units = "in", res = 600)
fig3
invisible(dev.off())
```




## Fig. 4: Priors vs Likelihood

### Prior, lik, post for current sample

```{r}
oll <- function(a1, x, y, a0, usd) {
  log_lik_vec <- dnorm(x = y, mean = (a0 + a1 * x), sd = usd, log = TRUE)
  sum_loglik <- sum(log_lik_vec)
  return(sum_loglik)
}

ols_model_small <- lm(y ~ x, data = sample1)
a1_postden <- approxfun(density(rescaled_posterior_2$a1))

small_sample_data <- tibble(a1 = seq(-20, +30, length.out = 1000))
small_sample_data$loglik <- map_dbl(
  small_sample_data$a1,
  oll,
  x = sample1$x,
  y = sample1$y,
  a0 = ols_model_small$coefficients[["(Intercept)"]],
  usd = sd(ols_model_small$residuals)
)
small_sample_data <-
  small_sample_data %>%
  mutate(
    avg_lik = exp(loglik / n_obs),
    prior = dnorm(a1, 0, 0.2 * (s_y / s_x)),
    post = a1_postden(a1)
  ) %>%
  select(-loglik)
small_sample_data$post[is.na(small_sample_data$post)] <- 0
```

### Panel A

```{r fig.height=3.5, fig.width=3.25, dpi=300}
a1_hat <- ols_model_small$coefficients[["x"]]
SCALER_1 <- (mean(small_sample_data$post) / mean(small_sample_data$avg_lik)) * 4
panA.4 <-
  small_sample_data %>%
  mutate(avg_lik = avg_lik * SCALER_1) %>%
  pivot_longer(c(avg_lik, prior, post), names_to = "part") %>%
  ggplot(aes(x = a1, y = value, group = part, color = part, fill = part)) +
  geom_ribbon(aes(ymin = 0, ymax = value), alpha = 0.5) +
  geom_line() +
  scale_color_manual(values = c("grey80", "grey30", "grey50")) +
  scale_fill_manual(values = c("grey80", "grey30", "grey50")) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    limits = c(0, 0.31),
    sec.axis = sec_axis(~ . / SCALER_1, name = " Likelihood (scaled by n)")
  ) +
  scale_x_continuous(
    breaks = c(-20, -10, 0, 2, 10, 20, 30),
    expand = expansion(mult = c(0, 0))
  ) +
  geom_vline(xintercept = a1_hat, color = "grey80", linetype = "dashed") +
  labs(
    y = "Density",
    x = expression(a[1]),
    subtitle = "Small sample (n = 50)"
  ) +
  annotate("text",
    x = c(-10, -4, 17),
    y = c(0.1, 0.2, 0.16),
    label = c("Prior", "Posterior", "Likelihood"),
    hjust = 0.5,
    size = 3
  ) +
  annotate("segment",
    x = c(-10, -4, 17),
    y = c(0.090, 0.189, 0.15),
    xend = c(-3.8, 2.5, 10),
    yend = c(0.050, 0.13, 0.09),
    color = c("grey50", "grey30", "grey80")
  ) +
  annotate("text",
    x = a1_hat + 1,
    y = 0.28, label = paste0("Max likelihood (", round(a1_hat, 1), ")"),
    hjust = 0, color = "grey50", size = 3
  ) +
  annotate("segment", x = 2, y = 0, xend = 2, yend = 0.005) +
  theme(legend.position = "none")
panA.4
```


### Gen larger sample

```{r}
set.seed(888)
n_obs2 <- 200
x <- rnorm(n = n_obs2, 0, 1)
sample2 <- gen_data(n_obs2)

input_data2 <- list(
  y = scale(sample2$y)[, 1],
  x = scale(sample2$x)[, 1],
  N = nrow(sample2)
)

u_y2 <- mean(sample2$y)
u_x2 <- mean(sample2$x)
s_y2 <- sd(sample2$y)
s_x2 <- sd(sample2$x)
```

### Fitting informative model to larger sample

```{r}
fit_3 <- stan(
  model_code = weakly_informative,
  data = input_data2,
  iter = 5000,
  warmup = 2000,
  chains = 2,
  seed = 1234
)

posterior_draws_3 <- as.data.frame(as.matrix(fit_3))
print(fit_3, pars = c("sigma", "a0", "a1"))
```

```{r}
rescaled_posterior_3 <- data.frame(
  a0 = posterior_draws_3$a0 * s_y2 + u_y2 - (posterior_draws_3$a1 * (s_y2 / s_x2) * u_x2),
  a1 = posterior_draws_3$a1 * (s_y2 / s_x2)
)

summary(rescaled_posterior_3)
```


### Prior, lik, post for larger sample

```{r}
oll <- function(a1, x, y, a0, usd) {
  log_lik_vec <- dnorm(x = y, mean = (a0 + a1 * x), sd = usd, log = TRUE)
  sum_loglik <- sum(log_lik_vec)
  return(sum_loglik)
}

ols_model_large <- lm(y ~ x, data = sample2)
a1_postden2 <- approxfun(density(rescaled_posterior_3$a1))

large_sample_data <- tibble(a1 = seq(-20, +30, length.out = 1000))
large_sample_data$loglik <- map_dbl(
  large_sample_data$a1,
  oll,
  x = sample2$x,
  y = sample2$y,
  a0 = ols_model_large$coefficients[["(Intercept)"]],
  usd = sd(ols_model_large$residuals)
)
large_sample_data <-
  large_sample_data %>%
  mutate(
    avg_lik = exp(loglik / n_obs2),
    prior = dnorm(a1, 0, 0.2 * (s_y2 / s_x2)),
    post = a1_postden2(a1)
  ) %>%
  select(-loglik)
large_sample_data$post[is.na(large_sample_data$post)] <- 0
head(large_sample_data)
```


### Panel B

```{r fig.height=3.5, fig.width=3.25, dpi=300}
a1_hat_big <- ols_model_large$coefficients[["x"]]
SCALER_2 <- (mean(large_sample_data$post) / mean(large_sample_data$avg_lik)) * 4
panB.4 <-
  large_sample_data %>%
  mutate(avg_lik = avg_lik * SCALER_2) %>%
  pivot_longer(c(avg_lik, prior, post), names_to = "part") %>%
  ggplot(aes(x = a1, y = value, group = part, color = part, fill = part)) +
  geom_ribbon(aes(ymin = 0, ymax = value), alpha = 0.5) +
  geom_line() +
  scale_color_manual(values = c("grey80", "grey30", "grey50")) +
  scale_fill_manual(values = c("grey80", "grey30", "grey50")) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    limits = c(0, 0.31),
    sec.axis = sec_axis(~ . / SCALER_2, name = " Likelihood (scaled by n)")
  ) +
  scale_x_continuous(
    breaks = c(-20, -10, 0, 2, 10, 20, 30),
    expand = expansion(mult = c(0, 0))
  ) +
  geom_vline(xintercept = a1_hat_big, color = "grey80", linetype = "dashed") +
  labs(
    y = "Density",
    x = expression(a[1]),
    subtitle = "Large Sample (n = 200)"
  ) +
  annotate("text",
    x = c(-10, -6, 17),
    y = c(0.1, 0.2, 0.16),
    label = c("Prior", "Posterior", "Likelihood"),
    hjust = 0.5,
    size = 3
  ) +
  annotate("segment",
    x = c(-10, -6, 17),
    y = c(0.090, 0.189, 0.15),
    xend = c(-3.8, 1.5, 10),
    yend = c(0.050, 0.13, 0.09),
    color = c("grey50", "grey50", "grey80")
  ) +
  annotate("text",
    x = a1_hat_big + 1,
    y = 0.28, label = paste0("Max likelihood (", round(a1_hat_big, 1), ")"),
    hjust = 0, color = "grey50", size = 3
  ) +
  annotate("segment", x = 2, y = 0, xend = 2, yend = 0.005) +
  theme(legend.position = "none")
panB.4
```

### Full figure

```{r fig.height=3.5, fig.width=6.5, dpi=300}
fig4 <- panA.4 + panB.4 +
  plot_annotation(tag_levels = "A")
fig4
```
### Saving figure

```{r fig.height=3.5, fig.width=6.5, dpi=600}
agg_png("../out/figs/fig4.png", width = 6.5, height = 3.5, units = "in", res = 600)
fig4
invisible(dev.off())
```


## OLS

```{r}
summary(lm(y ~ x, data = sample1))
```

```{r}
summary(lm(y ~ x, data = sample2))
```


## Fig 7. MCMC chains

### Individual chain draws inlcuding warmup phase 

```{r}
full_chains <- extract(fit_2, permuted = FALSE, inc_warmup = TRUE)

full_chains <-
  rbind(
    as.data.frame(full_chains[, 1, c("a0", "a1", "sigma")]) %>%
      mutate(
        chain = 1,
        step = 1:dim(full_chains)[1]
      ),
    as.data.frame(full_chains[, 2, c("a0", "a1", "sigma")]) %>%
      mutate(
        chain = 2,
        step = 1:dim(full_chains)[1]
      )
  ) %>%
  mutate(
    a0 = a0 * s_y2 + u_y2 - (a1 * (s_y2 / s_x2) * u_x2),
    a1 = a1 * (s_y2 / s_x2),
    chain = as.factor(chain)
  )
graph_annotate <- 
  full_chains %>% 
  select(-sigma) %>% 
  filter(step > 3000, step <= 3050)
```

### Panel A code

```{r  fig.height=3.5, fig.width=3.25, dpi=300}
panA.7 <- 
  full_chains %>% 
  filter(step <= 1020) %>% 
  pivot_longer(c(a0, a1, sigma), names_to = "pars") %>% 
  mutate(pars2 = case_when(
    pars == "a0" ~ "a[0]",
    pars == "a1" ~ "a[1]",
    pars == "sigma" ~ "sigma"
  )) %>% 
  ggplot(aes(x = step, y = value, group = chain, color = chain)) + 
  geom_line(alpha = 1, size = 0.5) + 
  facet_wrap(~pars, ncol = 1, scales = "free_y", labeller = label_parsed) + 
  scale_color_manual(values = c("gray10", "gray70")) +
  scale_x_continuous(expand = expansion(mult = c(0.02, 0))) +
  labs(
    subtitle = "Warmup phase of two MCMC chains",
    x = "Iteration", 
    y = NULL
    ) + 
  theme(legend.position = "none",
        strip.background = element_blank())
panA.7
```


### (not used)

```{r  fig.height=3.5, fig.width=3.25, dpi=300}
panB.7 <-
  ggplot(data = rescaled_posterior_2, aes(x = a0, y = a1)) +
  stat_density_2d(aes(fill = ..level..), geom = "polygon", colour = "white") +
  scale_x_continuous(expand = c(0, 0), limits = range_a0) +
  scale_y_continuous(expand = c(0, 0), limits = range_a1) +
  scale_fill_continuous(low = "grey95", high = "grey30") +
  theme(legend.position = "none") +
  labs(
    x = expression(a[0]),
    y = expression(a[1]),
    subtitle = quote("50 iterations of two MCMC chains")
  ) +
  geom_line(data = graph_annotate, 
            aes(x = a0, y = a1, group = chain, linetype = chain)) +
  geom_point(data = graph_annotate, 
             aes(x = a0, y = a1, group = chain, shape = chain)) +
  annotate("segment",
    x = c(0.30659650, 8.65481506),
    y = c(9.74364523, 4.74295103),
    xend = c(-5, 10),
    yend = c(12, 8),
    color = "grey60"
  ) +
  annotate("text", 
    x = c(-5, 10), 
    y = c(12.6, 8.6), 
    label = c("Chain 1", "Chain2"), 
    size = 3
  )
panB.7
```

### Panel B code

```{r fig.height=3.5, fig.width=3.25, dpi=300}
panC.7 <-
  rescaled_posterior_2 %>%
  ggplot(aes(x = a1)) +
  geom_histogram(binwidth = 0.3) +
  labs(
    x = expression(a[1]),
    y = expression("Number of samples"),
    subtitle = quote("6000 MCMC samples from posterior")
  ) +
  scale_x_continuous(expand = c(0, 0), limits = range_a1) +
  scale_y_continuous(expand = expansion(mult = c(0.0, 0.05)))
panC.7
```

### Panel C code

```{r  fig.height=3.5, fig.width=3.25, dpi=300}
panD.7 <-
  rescaled_posterior_2 %>%
  ggplot(aes(x = a0)) +
  geom_histogram(binwidth = 0.3) +
  labs(
    x = expression(a[0]),
    y = expression("Number of samples"),
    subtitle = quote("6000 MCMC samples from posterior")
  ) +
  scale_x_continuous(expand = c(0, 0), limits = range_a0) +
  scale_y_continuous(expand = expansion(mult = c(0.0, 0.05)))
panD.7
```

### Full figure

```{r  fig.height=7, fig.width=6.5, dpi=300}
layout <- "
AA
AA
BC
BD
"

fig7 <- panA.7 + panB.7 + panC.7 + panD.7 +
  plot_layout(design = layout) +
  plot_annotation(tag_levels = "A")
fig7
```


```{r  fig.height=3.5, fig.width=6.5, dpi=300}
layout <- "
AB
AC
"

fig7 <- panA.7 + panC.7 + panD.7 +
  plot_layout(design = layout) +
  plot_annotation(tag_levels = "A")
fig7
```

### Saving figure

```{r fig.height=3.5, fig.width=6.5, dpi=600}
agg_png("../out/figs/fig7.png", width = 6.5, height = 3.5, units = "in", res = 600)
fig7
invisible(dev.off())
```
