---
title: "Simulation: Classical vs. Bayesian Regression"
description: |
  Code for the Section 2: Bayesian inference compared to classical inference
author:
  - name: Harm Schuett
    url: https://hschuett.github.io/
    affiliation: Tilburg University
    affiliation_url: https://www.tilburguniversity.edu
date: "`r Sys.Date()`"
output: 
  distill::distill_article:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r xaringanExtra-clipboard, echo=FALSE}
xaringanExtra::use_clipboard()
```

```{r, message=FALSE}
library(tidyverse)
library(broom)
library(patchwork)
library(cmdstanr)
source("../R/00-utils.R")
kable <- knitr::kable
theme_set(theme_prodgray())
```

## 1. Introduction and preliminaries

This markdown file contains all the code necessary to replicate the figures, models and results used in Section 2: *Bayesian inference compared to classical inference* of the Paper *What Can Bayesian Inference Do for Accounting Research?*. All the code can also be found in the [repo](https://github.com/hschuett/BayesForAccountingResearch). It contains `00-utils.R` which contains a few helper functions for graphs and tables.

**Note:** I started using the newer [cmdstanr](https://mc-stan.org/cmdstanr/) package instead of the older [rstan](https://mc-stan.org/rstan/) package because it likely is the future of the R based Stan ecosystem. I also really like its api, which is very close to the api of the [pystan](https://pystan.readthedocs.io/en/latest/) package. An additional advantage (I hope) is thus that most model fitting code should be more or less directly transferable to pystan for those that want to work in python.

Installing cmdstanr used to be tricky at times because one needs a working c++ toolchain. But it is much smoother now. Please see the cmdstanr [doc](https://mc-stan.org/cmdstanr/reference/cmdstanr-package.html) for installation instructions

## 2. Creating the simulated data

First we create the 50 samples of 50 observations each

```{r}
set.seed(888)
n_samples <- 50
n_obs <- 50
x_fix <- rnorm(n = n_obs, 0, 1)
gen_data <- function(n) tibble(u = rnorm(n, 0, 20), x = x_fix, y = 1 + 2 * x_fix + u)

# gen samples
samples <- tibble(id = 1:n_samples)
samples$data <- map(rep.int(n_obs, n_samples), gen_data)
```

Next we create the relevant figures

## 3. OLS regressions

### 3.1. OLS fits of all 50 samples

```{r}
# gen ols estimates
samples$ols_fit <- map(samples$data, function(dat) tidy(lm(y ~ x, data = dat)))
ols_estimates <- samples %>%
  unnest(ols_fit) %>%
  mutate(param = if_else(term == "x", "a1", "a0")) %>%
  select(id, param, estimate) %>%
  spread(key = param, value = estimate) %>%
  mutate(
    plabs = paste0("(", round(a0, 1), ",", round(a1, 1), ")"),
    first_sample = as.factor(c(1, rep.int(0, times = n_samples - 1)))
  )
```

### 3.2. OlS estimates of the first sample

```{r}
kable(samples$ols_fit[[1]])
```

## 4. Fig. 1: Classical hypothesis tests

We need the following variables for scaling Fig. 1

```{r}
range_x <- range(ols_estimates$a0)
range_y <- range(ols_estimates$a1)
nudge_x <- 0.1
sample_a0 <- ols_estimates$a0[1]
sample_a1 <- ols_estimates$a1[1]
r <- sd(ols_estimates$a1)
```

### 4.1. Fig.1, Panel A

```{r fig.height=3.5, fig.width=3.25, dpi=150}
f1.panA <-
  ols_estimates %>%
  ggplot(aes(x = a0, y = a1, shape = first_sample, color = first_sample)) +
  geom_point(size = 2) +
  xlim(range_x) +
  ylim(range_y) +
  scale_color_manual(values = c("gray80", "black")) +
  labs(
    x = expression(" Intercept estimate " ~ hat(a)[0]),
    y = expression(" Slope estimate " ~ hat(a)[1]),
    subtitle = "Sampling variation"
  ) +
  geom_segment(aes(color = first_sample), xend = 1, yend = 2, linetype = 2, alpha = 0.5) +
  annotate("point", x = 1, y = 2, color = "black", size = 2, shape = 15) +
  annotate("text", x = 1.4, y = 1.9, label = "Truth (1, 2)", hjust = 0, color = "black", size = 3) +
  annotate("text",
    x = sample_a0 + 0.4,
    y = sample_a1 + -0.1,
    label = paste0("Sample (", round(sample_a0, 1), ", ", round(sample_a1, 1), ")"),
    hjust = 0, color = "black", size = 3
  ) +
  annotate("point", x = 0, y = 0, color = "black", size = 2, shape = 15) +
  annotate("text", x = 0.4, y = -0.1, label = "H0 (0, 0)", hjust = 0, color = "black", size = 3) +
  theme(legend.position = "none", aspect.ratio=1)
```

### 4.2. Fig.1, Panel B

```{r fig.height=3.5, fig.width=3.25, dpi=150}

f1.panB <- 
  ols_estimates %>%
  ggplot(aes(x = a0, y = a1, shape = first_sample, color = first_sample)) +
  geom_point(size = 2) +
  xlim(range_x) +
  ylim(range_y) +
  scale_color_manual(values = c("gray80", "black")) +
  labs(
    x = expression(" Intercept estimate " ~ hat(a)[0]),
    y = expression(" Slope estimate " ~ hat(a)[1]),
    subtitle = "Distance from hypothesis"
  ) +
  theme(legend.position = "none") +
  geom_segment(aes(color = first_sample), xend = 0, yend = 0, linetype = 2, alpha = 0.5) +
  annotate("point", x = 0, y = 0, color = "black", size = 2, shape = 15) +
  annotate("text", x = 0.4, y = -0.1, label = "H0 (0, 0)", hjust = 0, color = "black", size = 3) +
  annotate("text",
    x = sample_a0 + 0.4,
    y = sample_a1 + -0.1,
    label = paste0("Sample (", round(sample_a0, 1), ", ", round(sample_a1, 1), ")"),
    hjust = 0, color = "black", size = 3
  ) +
  geom_hline(yintercept = c(r, -r), color = "black") +
  geom_segment(
    x = -4, xend = -4, y = -r, yend = r, color = "black",
    arrow = arrow(length = unit(0.07, "inches"), ends = "both", type = "closed")
  ) +
  annotate("text",
    x = -3.4, y = -2.2,
    size = 3, hjust = 0, vjust = 0, color = "black",
    label = "One standard deviation"
  ) +
  theme(legend.position = "none", aspect.ratio=1)
```

Saving figure

```{r}
fig1 <- f1.panA + f1.panB + plot_annotation(tag_levels = "A")
save_fig(fig1, figname = "fig1", w = 6.2, h = 3.3)
```

```{r fig.height=3.5, fig.width=6.5, dpi=120}
fig1
```

## 5. Additional t-test figure

This one is not in the paper. I used it for an internal Ph.D. seminar

```{r fig.height=4, fig.width=6}
t_dist <- data.frame(ps = rt(n=10000, df=n_obs-2))
test_stat <- 1.65
quant_test <- round(1 - pt(test_stat, df=n_obs-2), 2)

fZ <- 
  ggplot(data=t_dist, aes(x=ps)) +
  geom_histogram(bins=50, color="white", fill="gray30") +
  geom_segment(color="red",
               x=test_stat, y=0,
               xend=test_stat, yend=200) +
  geom_segment(color="red",
               x=-test_stat, y=0,
               xend=-test_stat, yend=200) +
  # geom_area(color=high_red, fill=high_red, alpha=0.5) +
  annotate("text", x=test_stat, y=200,
           size=3, hjust=0, vjust=0,
           label=paste0("P(t >= ", test_stat, " | H0) = ", quant_test)) +
  annotate("text", x=-test_stat, y=200,
           size=3, hjust=1, vjust=0,
           label=paste0("P(t <= -", test_stat, " | H0) = ", quant_test)) +
  scale_y_continuous(expand=c(0,0)) +
  labs(x = expression((hat(a)-H0)/SE),
       y = NULL,
       subtitle = "Occurrence in repeated sampling") +
  theme(legend.position = "none")
fZ
```

## 6. Fitting the Bayesian models

To fit a Bayesian model I use [Stan](https://mc-stan.org/), or, more precisely, its R bindings in cmdstanr. To fit a Bayesian model, we need to:

1.  Write the corresponding model using the Stan language
2.  Compile the code into a Stan model executable (an .exe file)
3.  Make a list of data to feed into the .exe file
4.  Let the model run and generate MCMC chains (or whatever algorithm is specified)

### 6.1 Bayesian model with diffuse/very weak priors

```{r}
sample1 <- samples$data[[1]]
```

The following descriptive are useful when assessing what priors to use:

```{r}
sample1_descs <- c(
  u_y = mean(sample1$y),
  u_x = mean(sample1$x),
  s_y = sd(sample1$y),
  s_x = sd(sample1$x)
) 

round(sample1_descs, 3)
```

The model itself is coded in the [Stan language](https://mc-stan.org/). There are many excellent tutorials on Stan available online. So I won't waste space explaining it here. For various reasons (e.g., debugging) it is customary to put the model code in a separate .stan file. All the model files can be found in the /Stan/ folder of the [repo](https://github.com/hschuett/BayesForAccountingResearch).

```{r}
cat(read_lines("../Stan/sim-vweak-priors.stan"), sep = "\n")
```

<aside>

The model code in Stan language

</aside>

Next, we compile the model to an .exe file

```{r}
model_vweak_priors <- cmdstan_model("../Stan/sim-vweak-priors.stan")
```

Now, we prepare the list of data to feed into the model.

```{r}
input_data <- list(
  y = sample1$y,
  x = sample1$x,
  N = nrow(sample1)
)
```

<aside>

Note how the list matches the names of the variables declared in the model's data block

</aside>

We run the model

```{r}
fit_vweak_priors <- model_vweak_priors$sample(
  data = input_data,
  iter_sampling = 1000,
  iter_warmup = 1000,
  chains = 4,
  parallel_chains = 4,
  seed = 1234,
  refresh = 1000
)
```

Here is the summary of the resulting posterior distribution of the model parameters

```{r}
fit_vweak_priors$summary(variables = c("sigma", "a0", "a1")) |> 
  mutate(across(where(is.numeric), round, 3)) |> 
  kable()
```

#### Posterior descriptives

This is one form of getting the mcmc chain output. We need the draws to draw inferences about the posterior.

```{r}
postdraws_vweak_prior <- fit_vweak_priors$draws(format = "draws_df")
head(postdraws_vweak_prior)
```

The posterior 90% centered credible interval:

```{r}
quantile(postdraws_vweak_prior$a1, probs = c(0.05, 0.95))
```

The posterior probability that a1 \< 0:

```{r}
emp_a1 <- ecdf(postdraws_vweak_prior$a1)
emp_a1(0)
```

### 6.2. Bayesian model with weakly informative priors

This is basically the same model, just with different hard-coded priors.

```{r}
cat(read_lines("../Stan/sim-wkinfo-priors.stan"), sep = "\n")
```

```{r}
model_wkinfo_priors <- cmdstan_model("../Stan/sim-wkinfo-priors.stan")
```

Because we use the same data (The list `input_data`), we can fit the new model now:

```{r}
fit_wkinfo_priors <- model_wkinfo_priors$sample(
  data = input_data,
  iter_sampling = 1000,
  iter_warmup = 1000,
  chains = 4,
  parallel_chains = 4,
  seed = 1234,
  refresh = 1000
)
```

Here is the summary of the resulting posterior distribution of the model parameters

```{r}
fit_wkinfo_priors$summary(variables = c("sigma", "a0", "a1")) |> 
  mutate(across(where(is.numeric), round, 3)) |> 
  kable()
```

#### Posterior descriptives

The posterior 90% centered credible interval:

```{r}
postdraws_wkinfo_prior <- fit_wkinfo_priors$draws(format = "draws_df")
quantile(postdraws_wkinfo_prior$a1, probs = c(0.05, 0.95))
```

The posterior probability that a1 \< 0:

```{r}
emp_a1 <- ecdf(postdraws_wkinfo_prior$a1)
emp_a1(0)
```

## 7. Fig. 2: Visualization of Bayesian inference

### 7.1. Data preparations

The following code generates the density estimates, for the a1 posterior panels

```{r}
# helper function for panel D
get_dens <- function(ppd, probs = c(0.05, 0.95)) {
  dens <- density(ppd)
  mode3 <- dens$x[dens$y == max(dens$y)]
  mode3.den <- dens$y[dens$x == mode3]
  dens.df <- data.frame(x = dens$x, y = dens$y)
  quantiles <- quantile(ppd, prob = probs)
  dens.df$quant <- factor(findInterval(dens.df$x, quantiles))
  dens.df$Q1 <- quantiles[1]
  dens.df$Q2 <- quantiles[2]
  return(dens.df)
}

post_dens <- get_dens(postdraws_vweak_prior$a1)
post_dens2 <- get_dens(postdraws_wkinfo_prior$a1)

max_dens <- max(c(post_dens$y, post_dens2$y))
```

The following code generates the density estimates, for the prior panels

```{r}
x_length <- seq(-300, 300, length.out = 1000)
x_length2 <- seq(0, 100, length.out = 1000)
priors <- data.frame(
  a_den = dnorm(x = x_length, mean = 0, sd = 100),
  sigma_den = dexp(x = x_length2, rate = (1/21)),
  x1 = x_length,
  x2 = x_length2
)
priors2 <- data.frame(
  a_den = dnorm(x = x_length, mean = 0, sd = 4),
  sigma_den = dexp(x = x_length2, rate = (1/21)),
  x1 = x_length,
  x2 = x_length2
)

max_dens_a_prior <- max(c(priors$a_den, priors2$a_den))
```

Getting the plot ranges

```{r}
range_a0 <- c(
  min(min(postdraws_vweak_prior$a0), min(postdraws_vweak_prior$a0)), 
  max(max(postdraws_vweak_prior$a0), max(postdraws_vweak_prior$a0))
)
range_a1 <- c(
  min(min(postdraws_vweak_prior$a1), min(postdraws_vweak_prior$a1)), 
  max(max(postdraws_vweak_prior$a1), max(postdraws_vweak_prior$a1))
)
```

### 7.2. Fig.2, Subfigure 1

```{r}
f2.s1.panC <-
  ggplot(data = postdraws_vweak_prior, aes(x = a0, y = a1)) +
  stat_density_2d(aes(fill = ..level..), geom = "polygon", colour = "white") +
  annotate("point", x = 1, y = 2, color = "black", size = 2, shape = 15) +
  annotate("text", x = 1.4, y = 1.9, label = "Truth (1, 2)", hjust = 0, color = "black", size = 3) +
  annotate("point",
    x = mean(postdraws_vweak_prior$a0), y = mean(postdraws_vweak_prior$a1),
    color = "black", size = 2, shape = 17
  ) +
  annotate("text",
    x = -5, y = 10,
    label = paste0(
      "Posterior Mean\n(",
      round(mean(postdraws_vweak_prior$a0), 1),
      ", ",
      round(mean(postdraws_vweak_prior$a1), 1),
      ")"
    ),
    hjust = 0.5, color = "black", size = 3
  ) +
  annotate("segment",
    x = mean(postdraws_vweak_prior$a0), xend = -5,
    y = mean(postdraws_vweak_prior$a1), yend = 8.5
  ) +
  annotate("point", x = 0, y = 0, color = "black", size = 2, shape = 15) +
  annotate("text", x = 0.4, y = -0.1, label = "H0 (0, 0)", hjust = 0, color = "black", size = 3) +
  scale_x_continuous(expand = c(0, 0), limits = range_a0) +
  scale_y_continuous(expand = c(0, 0), limits = range_a1) +
  scale_fill_continuous(low = "grey95", high = "grey30") +
  theme(legend.position = "none") +
  labs(
    x = expression(a[0]),
    y = expression(a[1]),
    subtitle = quote("Posterior: Pr(" * a[0] * "," * a[1] * " | y, x)")
  )
f2.s1.panA <-
  ggplot(data = priors, aes(x = x1, y = a_den)) +
  geom_area(color = "black", fill = "grey30", alpha = 0.5) +
  scale_y_continuous(expand = expansion(mult = c(0.001, 0.05)), limits = c(0, max_dens_a_prior) ) +
  annotate("text", x = 290, y = 0.095, label = "N(0, 100)", hjust = 1, color = "black", size = 3) +
  scale_x_continuous(expand = c(0, 0), breaks = c(-300, -150, 0, 150, 300)) +
  labs(
    y = "Density",
    x = expression(a[k]),
    subtitle = quote("Prior: Pr(" * a[k] * ")")
  )
f2.s1.panB <-
  ggplot(data = priors, aes(x = x2, y = sigma_den)) +
  geom_area(color = "black", fill = "grey30", alpha = 0.5) +
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.05))) +
  scale_x_continuous(expand = c(0, 0)) +
  annotate("text", x = 99, y = 0.045, label = "Exp(1/21)", hjust = 1, color = "black", size = 3) +
  labs(
    y = "Density",
    x = expression(sigma),
    subtitle = quote("Prior: Pr(" * sigma * ")")
  )
f2.s1.panD <-
  post_dens %>%
  ggplot(aes(x = x)) +
  geom_ribbon(aes(ymin = 0, ymax = y, fill = quant), alpha = 0.7) +
  geom_line(aes(y = y), color = "black") +
  labs(
    x = expression(a[1]),
    y = expression("Density"),
    subtitle = quote("Posterior: Pr(" * a[1] * " | y, x)")
  ) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = expansion(mult = c(0.0, 0.05))) +
  scale_fill_manual(values = c("grey50", "grey70", "grey50")) +
  theme(legend.position = "None") +
  annotate("segment", x = 7, xend = 12, y = 0.02, yend = 0.05) +
  annotate("text", x = 11.5, y = 0.052, label = "90% Credible\ninterval", color = "black", size = 3, hjust = 0) +
  coord_flip(xlim = range_a1, ylim = c(0, max_dens))

layout <- "
ACCD
BCCD
"

fig2.s1 <- f2.s1.panA + f2.s1.panB + f2.s1.panC + f2.s1.panD +
  plot_layout(design = layout) +
  plot_annotation(tag_levels = "A")
```

### 7.3. Fig.2, Subfigure 2

```{r}
f2.s2.panC <-
  ggplot(data = postdraws_wkinfo_prior, aes(x = a0, y = a1)) +
  stat_density_2d(aes(fill = ..level..), geom = "polygon", colour = "white") +
  annotate("point", x = 1, y = 2, color = "black", size = 2, shape = 15) +
  annotate("text", x = 1.4, y = 1.9, label = "Truth (1, 2)", hjust = 0, color = "black", size = 3) +
  annotate("point",
    x = mean(postdraws_wkinfo_prior$a0), y = mean(postdraws_wkinfo_prior$a1),
    color = "black", size = 2, shape = 17
  ) +
  annotate("text",
    x = -5, y = 10,
    label = paste0(
      "Posterior Mean\n(",
      round(mean(postdraws_wkinfo_prior$a0), 1),
      ", ",
      round(mean(postdraws_wkinfo_prior$a1), 1),
      ")"
    ),
    hjust = 0.5, color = "black", size = 3
  ) +
  annotate("segment",
    x = mean(postdraws_wkinfo_prior$a0), xend = -5,
    y = mean(postdraws_wkinfo_prior$a1), yend = 8.5
  ) +
  annotate("point", x = 0, y = 0, color = "black", size = 2, shape = 15) +
  annotate("text", x = 0.4, y = -0.1, label = "H0 (0, 0)", hjust = 0, color = "black", size = 3) +
  scale_x_continuous(expand = c(0, 0), limits = range_a0) +
  scale_y_continuous(expand = c(0, 0), limits = range_a1) +
  scale_fill_continuous(low = "grey95", high = "grey30") +
  theme(legend.position = "none") +
  labs(
    x = expression(a[0]),
    y = expression(a[1]),
    subtitle = quote("Posterior: Pr(" * a[0] * "," * a[1] * " | y, x)")
  )
f2.s2.panA <-
  ggplot(data = priors2, aes(x = x1, y = a_den)) +
  geom_area(color = "black", fill = "grey30", alpha = 0.5) +
  scale_y_continuous(expand = expansion(mult = c(0.001, 0.05)), limits = c(0, max_dens_a_prior)) +
  annotate("text", x = 290, y = 0.095, label = "N(0, 4)", hjust = 1, color = "black", size = 3) +
  scale_x_continuous(expand = c(0, 0), breaks = c(-300, -150, 0, 150, 300)) +
  labs(
    y = "Density",
    x = expression(a[k]),
    subtitle = quote("Prior: Pr(" * a[1] * ")")
  )
f2.s2.panB <-
  ggplot(data = priors2, aes(x = x2, y = sigma_den)) +
  geom_area(color = "black", fill = "grey30", alpha = 0.5) +
  scale_y_continuous(expand = expansion(mult = c(0.01, 0.05))) +
  scale_x_continuous(expand = c(0, 0)) +
  annotate("text", x = 99, y = 0.045, label = "Exp(1/21)", hjust = 1, color = "black", size = 3) +
  labs(
    y = "Density",
    x = expression(sigma),
    subtitle = quote("Prior: Pr(" * sigma * ")")
  )
f2.s2.panD <-
  post_dens2 %>%
  ggplot(aes(x = x)) +
  geom_ribbon(aes(ymin = 0, ymax = y, fill = quant), alpha = 0.7) +
  geom_line(aes(y = y), color = "black") +
  labs(
    x = expression(a[1]),
    y = expression("Density"),
    subtitle = quote("Posterior: Pr(" * a[1] * " | y, x)")
  ) +
  scale_x_continuous(expand = c(0, 0), limits = range_a1) +
  scale_y_continuous(expand = expansion(mult = c(0.0, 0.05))) +
  scale_fill_manual(values = c("grey50", "grey70", "grey50")) +
  theme(legend.position = "None") +
  annotate("segment", x = 7, xend = 12, y = 0.02, yend = 0.05) +
  annotate("text", x = 11.5, y = 0.052, label = "90% Credible\ninterval", color = "black", size = 3, hjust = 0) +
  coord_flip(xlim = range_a1, ylim = c(0, max_dens))

layout <- "
ACCD
BCCD
"

fig2.s2 <- f2.s2.panA + f2.s2.panB + f2.s2.panC + f2.s2.panD +
  plot_layout(design = layout) +
  plot_annotation(tag_levels = "A")
```

```{r fig.height=3.5, fig.width=6.5, dpi=120}
fig2.s1
```

```{r fig.height=3.5, fig.width=6.5, dpi=120}
fig2.s2
```

Saving subfigures

```{r}
save_fig(fig=fig2.s1, figname = "fig2-1", w = 6.2, h = 3.3)
save_fig(fig=fig2.s2, figname = "fig2-2", w = 6.2, h = 3.3)
```

## 8. Fig. 3: Priors vs Likelihood

### 8.1. Generate a larger sample

```{r}
set.seed(999)
new_obs <- 150
x_fix2 <- rnorm(n = new_obs, 0, 1)
sample2 <- tibble(u = rnorm(new_obs, 0, 20), x = x_fix2, y = 1 + 2 * x_fix2 + u)
sample2 <- bind_rows(sample1, sample2)
n_obs2 <- new_obs + n_obs

input_data2 <- list(
  y = sample2$y,
  x = sample2$x,
  N = nrow(sample2)
)
```

### 8.2. Fitting the weakly informative model to larger sample

```{r}
fit_wkinfo_priors_large <- model_wkinfo_priors$sample(
  data = input_data2,
  iter_sampling = 1000,
  iter_warmup = 1000,
  chains = 4,
  parallel_chains = 4,
  seed = 1234,
  refresh = 1000
)
```

```{r}
postdraws_wkinfo_prior_large <- fit_wkinfo_priors_large$draws(format = "draws_df")
fit_wkinfo_priors_large$summary(variables = c("sigma", "a0", "a1")) |> 
  mutate(across(where(is.numeric), round, 3))  |> 
  kable()
```

### 8.3. Collecting Prior, likelihood, posterior for small, current sample

```{r}
loglik_a1 <- function(a1, x, y, a0, u_sd) {
  resids <- y - (a0 + a1 * x)
  log_lik_vec <- dnorm(x = resids, mean = 0, sd = u_sd, log = TRUE)
  log_lik <- sum(log_lik_vec)
  return(log_lik)
}
```

```{r}
ols_model_small <- lm(y ~ x, data = sample1)
a0 <- ols_model_small$coefficients[["(Intercept)"]]
usd <- sd(ols_model_small$residuals)
a1_hat <- ols_model_small$coefficients[["x"]]

small_sample_data <- tibble(a1 = seq(-20, +30, length.out = 1000))
small_sample_data$loglik <- map_dbl(
  small_sample_data$a1,
  loglik_a1,
  x = sample1$x,
  y = sample1$y,
  a0 = a0,
  u_sd = usd
)
small_sample_data <-
  small_sample_data %>%
  mutate(
    # norm_lik = -2 * loglik, #exp(loglik) / sum(exp(loglik)),
    norm_lik = (loglik),
    prior = dnorm(a1, 0, 4),
    post = approxfun(density(postdraws_wkinfo_prior$a1))(a1)
  ) 
small_sample_data$post[is.na(small_sample_data$post)] <- 0
```

### 8.4. Collecting Prior, likelihood, posterior for larger sample

```{r}
ols_model_large <- lm(y ~ x, data = sample2)
a0 <- ols_model_large$coefficients[["(Intercept)"]]
usd <- sd(ols_model_large$residuals)
a1_hat_big <- ols_model_large$coefficients[["x"]]

large_sample_data <- tibble(a1 = seq(-20, +30, length.out = 1000))
large_sample_data$loglik <- map_dbl(
  large_sample_data$a1,
  loglik_a1,
  x = sample2$x,
  y = sample2$y,
  a0 = a0,
  u_sd = usd
)
large_sample_data <-
  large_sample_data %>%
  mutate(
    # norm_lik = -2 * loglik, #exp(loglik) / sum(exp(loglik)),
    norm_lik = (loglik),
    prior = dnorm(a1, 0, 4),
    post = approxfun(density(postdraws_wkinfo_prior_large$a1))(a1)
  ) 
large_sample_data$post[is.na(large_sample_data$post)] <- 0
```

### 8.5. Fig. 3, Panel A and Panel B

```{r fig.height=3.5, fig.width=6.5, dpi=120}
SCALER_1 <- -2 * 0.0001
f3.panA <- 
  small_sample_data %>%
  mutate(norm_lik = (-1600 - norm_lik) * SCALER_1) %>%
  pivot_longer(c(norm_lik, prior, post), names_to = "part") %>%
  ggplot(aes(x = a1, y = value, group = part, color = part, fill = part)) +
  geom_ribbon(aes(ymin = 0, ymax = value), alpha = 0.5) +
  geom_line() + 
  geom_vline(xintercept = a1_hat, color = "grey80", linetype = "dashed") +
  labs(
    y = "Density",
    x = expression(a[1]),
    subtitle = "Small sample (n = 50)"
  ) +
  annotate("text",
    x = c(-10, -4, 17),
    y = c(0.1, 0.2, 0.20),
    label = c("Prior", "Posterior", "Log likelihood"),
    hjust = 0.5,
    size = 3
  ) +
  annotate("segment",
    x = c(-10, -4, 17),
    y = c(0.090, 0.189, 0.21),
    xend = c(-3.8, 2.5, 10),
    yend = c(0.050, 0.13, 0.27),
    color = c("grey50", "grey30", "grey80")
  ) +
  annotate("text",
    x = a1_hat + 0.1,
    y = 0.31, label = paste("Max likelihood at", round(a1_hat, 1)),
    hjust = 0, color = "grey50", size = 3
  )
  

f3.panB <- 
  large_sample_data %>%
  mutate(norm_lik = (-1600 - norm_lik) * SCALER_1) %>%
  pivot_longer(c(norm_lik, prior, post), names_to = "part") %>%
  ggplot(aes(x = a1, y = value, group = part, color = part, fill = part)) +
  geom_ribbon(aes(ymin = 0, ymax = value), alpha = 0.5) +
  geom_line() +
  geom_vline(xintercept = a1_hat_big, color = "grey80", linetype = "dashed") +
  labs(
    y = "Density",
    x = expression(a[1]),
    subtitle = "Large Sample (n = 200)"
  ) +
  annotate("text",
    x = c(-10, -6, 17),
    y = c(0.1, 0.2, 0.20),
    label = c("Prior", "Posterior", "Log likelihood"),
    hjust = 0.5,
    size = 3
  ) +
  annotate("segment",
    x = c(-10, -6, 17),
    y = c(0.090, 0.189, 0.19),
    xend = c(-3.8, 1.5, 10),
    yend = c(0.050, 0.13, 0.13),
    color = c("grey50", "grey50", "grey80")
  ) +
  annotate("text",
    x = a1_hat_big + .1,
    y = 0.31, label = paste("Max likelihood at", round(a1_hat_big, 1)),
    hjust = 0, color = "grey50", size = 3
  ) 

fig3 <- 
  f3.panA + f3.panB +
  plot_annotation(tag_levels = "A") &
  annotate("segment", x = 2, y = 0, xend = 2, yend = 0.02) &
  theme(legend.position = "none", aspect.ratio=1) & 
  scale_color_manual(values = c("grey80", "grey30", "grey50")) &
  scale_fill_manual(values = c("grey80", "grey30", "grey50")) &
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    limits = c(0, 0.32),
    sec.axis = sec_axis(~ -1600 - . / SCALER_1, name = "Log likelihood")
  ) &
  scale_x_continuous(
    breaks = c(-20, -10, 0, 2, 10, 20, 30),
    expand = expansion(mult = c(0, 0))
  )
fig3
```

Saving figure

```{r}
save_fig(fig=fig3, figname = "fig3", w = 6.2, h = 3.1)
```
